|

### 2021-03-26 paper/v2_taylor_francis/response_reviewers.md

---
output:
  pdf_document: default
bibliography: bibliography.bib
suppress-bibliography: true
---

@@

We have added the following sentence to the first paragraph in section 3.2, explaining
why we believe any bias balances out:
"Indeed, those who seek to test their broadband are most likely to do so because
they are experiencing slower speeds than expected, although any skew towards slower
speeds is balanced by the likelihood that those who test their broadband are also
more 'tech-savvy' and / or have purchased higher speed packages that are not
delivering the promised level of service."

As we explain, our intentions were to analyse a dataset of experienced internet speeds.
Those who use the speedcheck service are not solicited in any way, and no personally
identifying data is included other than the geolocation of the test.

> 2. The results for network speed for the different regions are not very different with 1-2 out of 10Mbps difference and probably can not reflect the social disparities. Not sure if this is the facts in U.K. If so, then it's not a problem?

We have no reasons to expect that mean speeds are that different between regions of the UK
(e.g. North or South England). The greater differences are between urban, built-up
areas, and more rural areas within regions, although there are some local initiatives
to address this in some areas. However, we assume that your comment refers to Table 1
and the differences in mean network speed between clusters? Whilst these differences
are not large, the range for upload speeds is smaller than for download speeds,
and a difference of 1-2Mbps slower may well be noticeable at the wrong moment.
Importantly, we only consider the tests which took place during working hours, so
when speed drops, let's say, during a work Zoom call or when uploading large files,
it will be noticeable. This is why we created the clusters by how the time series
of speeds changes by day of the week and hour of the day, rather than focusing our
analysis on mean speeds for the period as a whole.

@@

> 3. The relevance to pandemic is not very straightforward. This may be improved by adding some studies, e.g., how the clusters impact the effectiveness of work from home through survey. Or relates to the employment rate.

Regrettably, we did not have the opportunity to run a survey. 
As we say in the introduction, prior to the pandemic, the demand for telecommuting 
would not have put a strain on internet connection speeds. We have clarified this 
with the addition of the following sentence:
"High levels of demand are one of the main causes of reduced reliability and slower 
connection speeds, as network bandwidth becomes congested".

We also have added a variable to the auxiliary regression tracking the share of
the population that was put on furlough. This was the scheme in the UK which, at
least until now, prevented much change in the employment rate, but did show who
could work and who could not at the height of the first wave of the pandemic.

@@

"This finding broadly applies in the UK, where studies that also analysed broadband
speed checks concluded that average speeds are lower in rural areas, something that
has not been improved by policy measures to increase competition 
[@riddlesden2014broadband; @nardotto2015unbundling]. In contrast,
@riddlesden2014broadband found that levels of deprivation did not correlate with
first level digital divides".

and

"Dense urban areas were shown to suffer more from slow down during peak hours,
although these services were more likely to be improved by increased competition
between providers, such as between new entrants and Virgin Media cable connections
[@riddlesden2014broadband; @nardotto2015unbundling]. The latter were historically
available to only $45$% of premises in the UK [@ofcom2016], where the more
lucrative and competitive market originally attracted the cable TV provider".

The latter bit about Virgin Media is moved from the results section, in response to Reviewer 2, q7.

@@

> 1. The paper used k-means, a popular unsupervised classification in this study. The optimal number of clusters was determined as 13 based on cluster validation indices (CVIs). While this number could be well justified, I feel it is too high to properly interpret the results. The interpretation and knowledge discovery from clustering results with a high k tend to be tedious. Being affected by the high k, the Results section appears to be very difficult to read and follow. As there are many different CVIs, I feel a lower k will benefit the interpretation of results.

Please see our response to the next point.

> 2. Another major comment is on the multinomial logit regression (MLR). This is partly related to my first comment on a larger k. While there is no hard limit on the number of categories in MLR, the interpretation of results will be difficult with a high k as well. In addition, if the dependence among independent variables (collinearity) occurs, the estimation of the multinomial logit model parameters becomes inaccurate. According to Table 2, it is very likely some of the variables could present collinearity (e.g., distance to nearest metro and tech jobs).

Following your suggestion, we have now opted to present the cluster analysis and
the subsequent multinomial regression using $k = 9$. We indeed believe that the
results are clearer now. We have added the following justification in the Methods
and Data section:

@@

> 5.  There are three levels of digital divide. According to the last paragraph in Section 2, only the intersection of level-1 and level-3 was included in this research. I am wondering how level-2 may fit in the general research framework. Could education and demographic variables be used to represent level-2 digital divide?

We clarify that the occupations included in the regression are more about the level-2
digital divide, providing an indication of not only education, but digital
skills. The average earnings and, particularly during the pandemic, the new variable 
-- % furloughed, are indications of level-3 digital divides. We did have education
in the previous submission (NVQ4+), but have removed it due to multicollinearity
-- see response for point 2.

> 6. A few suggestions on the figures. Fig 1 and 2 have crowded tick labels for x-axis month and weekday. Fig 3 and 4 may be placed side by side with the same y-axis for easy comparison. Actually, I feel “Figure 3 shows that the high standard deviation” may be inaccurate as Fig 4 appears to exhibit higher variability. Fig 5 maps may include the locations of metropolitan or major cities to help interpretation. The warmer-cooler color scheme may match the population size instead of the ordinal cluster numbers. The current color scheme does not help interpretation at all.

- x-axis for Figures 1 and 2 have been amended accordingly.

- The temporal profiles of the upload clusters are now included in Figure 3.

- Figure 5, which is now labeled Figure 4: the map has now been redesigned using a different
colour palette to better reflect the discrete nature of the underpinning data --
the time-series clusters. Points and labels of the main UK cities were also added.

@@

> 7. A few minor things. In Sec 2.2, the statement “These returns have been even greater …” appears to be controversial. LAD acronym appears before it was defined. Virgin Media as a major broadband provider should be briefly introduced in the manuscript.

- We have toned down the statement on returns to say:
"In some aspects, these returns may have increased during the Covid-19 crisis..."

- Apologies on LAD acronym, now corrected.

- We also moved and integrated introduction of Virgin Media to the literature review s2.2. See also response to Reviewer 1, q6.

### 2021-03-26 paper/v2_taylor_francis/broadband.speed.covid.tex
second submission

\author{\name{Hannah Budnitz$^{a}$, Emmanouil Tranos$^{b, c}$}
\affil{$^{a}$Transport Studies Unit, University of Oxford, UK; $^{b}$School of Geographic Sciences, University of Bristol, UK; $^{c}$The Alan Turing Institute, UK}
}

\thanks{CONTACT Hannah Budnitz. Email: , Emmanouil Tranos. Email: \href{mailto:e.tranos@bristol.ac.uk}{\nolinkurl{e.tranos@bristol.ac.uk}}}

@@

Of the smaller clusters, cluster \(7\) experiences quite similar mean
speeds to cluster \(9\), and a not too dissimilar standard deviation,
but mean speeds in the morning peak are almost \(10\)\% lower than the
evening peak. Cluster \(3\) boasts speeds almost \(1\)Mb/s higher again,
but suffers from even greater slowdown in the morning peak. In
comparison, clusters \(2\) and \(5\), home to a little less than half a
million people, experience not only above average mean speeds, but also
faster speeds in the morning peak compared to the evening. Thus, in the
five LADs in these two clusters, the temporal profile of internet use
may be closer to what might have been expected pre-pandemic. In
comparison, the \(15\) LADs in clusters \(1\) and \(8\), home to over
\(1.4\) million people, not only experience fairly average upload speeds
and high ratios of standard deviation to the mean, but also experience
much lower speeds during the morning peak than the evening peak. From
the large spikes and dips shown on Figure \ref{UpCluster}, it is likely
that this poor reliability or consistency of morning internet speeds was
more noticeable to those in cluster \(1\), whilst those in cluster \(8\)
experienced the lack of reliable service as a problem throughout the
day.

It is worth noting here that clusters of LAD based on the upload speeds
are very different to the ones occur when we apply the same methods to
upload speed data for the same time period in 2019 -- see Appendix
\protect\hyperlink{appendix2}{2}. This drastic difference between the
LAD upload speed profiles before and during the pandemic is an
indication of the changes in internet usage that took place during the
pandemic.

>>>>

\author{\name{Hannah Budnitz$^{a}$, Emmanouil Tranos$^{b, c, \dagger}$}
\affil{$^{a}$Transport Studies Unit, University of Oxford, UK; $^{b}$School of Geographic Sciences, University of Bristol, UK; $^{c}$The Alan Turing Institute, UK; $^{\dagger}$Corresponding author}
}

\thanks{CONTACT Hannah Budnitz. Email: \href{mailto:hannah.budnitz@ouce.ox.ac.uk}{\nolinkurl{hannah.budnitz@ouce.ox.ac.uk}}, Emmanouil Tranos. Email: \href{mailto:e.tranos@bristol.ac.uk}{\nolinkurl{e.tranos@bristol.ac.uk}}}

@@

Of the smaller clusters, the \(15\) LADs in clusters \(1\) and \(8\),
home to over \(1.4\) million people, not only experience fairly average
upload speeds and high ratios of standard deviation to the mean, but
also experience much lower speeds during the morning peak than the
evening peak. From the large spikes and dips shown on Figure
\ref{UpCluster}, it is likely that this poor reliability or consistency
of morning internet speeds was more noticeable to those in cluster
\(1\), whilst those in cluster \(8\) experienced the lack of reliable
service as a problem throughout the day. Meanwhile, cluster \(7\)
experiences quite similar mean speeds to cluster \(9\), and a not too
dissimilar standard deviation, but mean speeds in the morning peak are
almost \(10\)\% lower than the evening peak. Cluster \(3\) boasts speeds
almost \(1\)Mb/s higher again, but suffers from even greater slowdown in
the morning peak. In comparison, clusters \(2\) and \(5\), home to a
little less than half a million people, experience not only above
average mean speeds, but also faster speeds in the morning peak compared
to the evening. Thus, in the five LADs in these two clusters, the
temporal profile of internet use may be closer to what might have been
expected pre-pandemic.

It is worth noting here that when we apply the same methods to upload
speed data for the same time period in 2019, the spatial distribution
and composition of the nine clusters of LADs are very different -- see
Appendix \protect\hyperlink{appendix2}{2}. This difference is an
indication of the changes in the temporal profile of internet usage that
took place during the pandemic, and which would be independent of any
wider trend in improving internet services.

### 2021-03-26 paper/v2_taylor_francis/broadband.speed.covid_anonymised.Rmd
second submission

---
title: |
  Working from home and digital divides: resilience during the pandemic
author:
  - name: Hannah Budnitz
    affil: a
  - name: Emmanouil Tranos
    affil: b, c
    email: e.tranos@bristol.ac.uk
affiliation:
  - num: a
    address: |
      Transport Studies Unit, University of Oxford, UK
  - num: b
    address: |
      School of Geographic Sciences, University of Bristol, UK
  - num: c
    address: |
      The Alan Turing Institute, UK
bibliography: bibliography.bib
abstract: |
  This paper offers a new perspective on telecommuting from the viewpoint of the complex web of digital divides. Using the UK as a case study, this paper studies how the quality and reliability of internet services, as reflected in *experienced* internet upload speeds during the spring 2020 lockdown, may reinforce or redress the spatial and social dimensions of digital divisions. Fast, reliable internet connections are necessary for the population to be able to work from home. Although not every place hosts individuals in occupations which allow for telecommuting nor with the necessary skills to effectively use the internet to telecommute, good internet connectivity is also essential to local economic resilience in a period like the current pandemic. Employing data on individual broadband speed tests and state-of-the-art time-series clustering methods, we create clusters of UK local authorities with similar temporal signatures of experienced upload speeds. We then associate these clusters of local authorities with their socioeconomic and geographic characteristics to explore how they overlap with or diverge from the existing economic and digital geography of the UK. Our analysis enables us to better understand how the spatial and social distribution of both occupations and online accessibility intersect to enable or hinder the practice of telecommuting at a time of extreme demand.
keywords: |
  covid; internet; working from home; broadband speed; time-series clusters
header-includes: |
  \usepackage{hyperref}
  \usepackage[utf8]{inputenc}
  \usepackage{dcolumn}
  \def\tightlist{}
output: rticles::tf_article
---

@@

>>>>

---
title: |
  Working from home and digital divides: resilience during the pandemic
author:
  - name: 
    affil: a
affiliation:
  - num: a
    address: 
bibliography: bibliography.bib
abstract: |
  This paper offers a new perspective on telecommuting from the viewpoint of the complex web of digital divides. Using the UK as a case study, this paper studies how the quality and reliability of internet services, as reflected in *experienced* internet upload speeds during the spring 2020 lockdown, may reinforce or redress the spatial and social dimensions of digital divisions. Fast, reliable internet connections are necessary for the population to be able to work from home. Although not every place hosts individuals in occupations which allow for telecommuting nor with the necessary skills to effectively use the internet to telecommute, good internet connectivity is also essential to local economic resilience in a period like the current pandemic. Employing data on individual broadband speed tests and state-of-the-art time-series clustering methods, we create clusters of UK local authorities with similar temporal signatures of experienced upload speeds. We then associate these clusters of local authorities with their socioeconomic and geographic characteristics to explore how they overlap with or diverge from the existing economic and digital geography of the UK. Our analysis enables us to better understand how the spatial and social distribution of both occupations and online accessibility intersect to enable or hinder the practice of telecommuting at a time of extreme demand.
keywords: |
  covid; internet; working from home; broadband speed; time-series clusters
header-includes: |
  \usepackage{hyperref}
  \usepackage[utf8]{inputenc}
  \usepackage{dcolumn}
  \def\tightlist{}
output: rticles::tf_article
---

@@

\newpage
